---
title: "FDA FAERS Data Extraction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(XML)
library(tidyverse)
```

# Note:

This file was the start of an attempt to create an easily reproducible means of accessing the FAERS database; thus potentially allowing a full PV analysis environment within R. This attempt was quickly abandoned due to improper HTML structure on the FDA website. Even if the HTML structure is fixed on a later date, a persistent issue with this implementation will be the size and uncleaned nature of the quarterly files.

### Accessing the data

The data for all FAERS per quarter per year is publicly available at https://www.fda.gov/Drugs/GuidanceComplianceRegulatoryInformation/Surveillance/AdverseDrugEffects/ucm083765.htm. To download these files, a web-crawler function has been written which reads the HTML content of this page to identify all links for a specified window of time.

```{r}
get_faers_urls <- function(url='https://www.fda.gov/Drugs/GuidanceComplianceRegulatoryInformation/Surveillance/AdverseDrugEffects/ucm083765.htm',
                           start_period, end_period, data_type='ASCII') {
  ## Crawls the HTML content of the FDA FAERS download page
  ## to extract URLs for downloading the requested data.
  ##
  ## Arguments:
  ## url:          URL to parse for data. Default is current page as of June 2017.
  ## start_period: a vector of year and quarter info to specify the first included data.
  ##               e.g.: c(2005, 1) <==> Jan. - March 2005
  ## end_period:   same as start_period, but for the last included data.
  ## data_type:    character; specify either 'ASCII' or 'SGML'.
  
  
  ## Check arguments
  if(!(data_type == 'ASCII' | data_type == 'SGML')) {
    stop('Data type must be either ASCII or SGML')
  }
  if(start_period[2] > 4 | start_period[2] < 1) {
    stop('Invalid quarter specification for start period')
  }
  if(end_period[2] > 4 | end_period[2] < 1) {
    stop('Invalid quarter specification for end period')
  }
  if(start_period[1] < 2004 | start_period[1] > 2011) {
    stop('Invalid start year: must be between 2004 - 2011')
  }
  if(end_period[1] < 2004 | end_period[1] > 2011) {
    stop('Invalid end year: must be between 2004 - 2011')
  }
  
  ## Create character strings for each quarter data
  year_sequence <- expand.grid(qtr = c(1:4), 
                               yr = seq(start_period[1], end_period[1])) %>% as.tbl
  
  selected_window <- year_sequence %>% filter(
    (yr > start_period[1] | qtr >= start_period[2]) &
    (yr < end_period[1] | qtr <= end_period[2]))
  
  selected_window_strings <- selected_window %>% 
    transmute(strings = 
    paste(as.character(yr), 'q', as.character(qtr), sep = '')) %>%
    .$strings

  ## Locate HTML content
  html_content <- htmlParse(url)
  
  
}
```

```{r}
html_content <- htmlParse(file = 'https://www.fda.gov/Drugs/GuidanceComplianceRegulatoryInformation/Surveillance/AdverseDrugEffects/ucm082193.htm')

## Unfortunately, stop point is here for now, because website HTML content is not being parsed correctly

## Could manually create list of download URLs but this would not be an efficient use of time
```

```{r}
html_content %>%
  xmlRoot %>% getNodeSet('//BODY')
```




```{r}
sessionInfo()
```

